The Efficient Handling of Guards in the Design of RPython's Tracing JIT

### 2012-08-16 talk/vmil2012/paper.tex
typo
calculate the percentage of guards responsible for 50% of the guard failures for the different benchmarks

It is important to handle guards well, because they are very common operations
in the traces produced by tracing JITs. As can be seen in
Figure~\ref{fig:guard_percent} guards account for about 14\% to 22\% of the
operations before and for about 15\% to 20\% of the operations after optimizing
the traces generated for the different benchmarks used in this paper. An
additional property is that guard failure rates are very uneven. The majority
of guards never fail at all, whereas those that do usually fail extremely
often.

Besides being common, guards have various costs associated to them.
Guards have a runtime cost, they take time to execute. Therefore it is
important to make the on-trace execution of guards as efficient as possible. On
the other hand, guards are possible deoptimization points. The recorded and
compiled path has to be left if a guard fails, returning control to the
interpreter. Therefore guards need enough associated information to enable
rebuilding the interpreter state. The memory overhead of this information
should be kept low. These constraints and trade-offs are what makes the design
and optimization of guards an important and non-trivial aspect of the construction
of a tracing just-in-time compiler.

@@

\begin{figure}
    \include{figures/guard_table}
    \caption{Percentage of guards before and after optimization for different benchmarks}
    \label{fig:guard_percent}
\end{figure}

@@

\begin{figure}
    \include{figures/failing_guards_table}
    \caption{Failing guards relative to the total number of guards}
    \label{fig:failing_guards}
\end{figure}

From Figure~\ref{fig:failing_guards} we can see that only a very small amount
of all the guards in the compiled traces ever fail. This amount varies between
2.4\% and 5.7\% of all guards. As can be expected, even fewer guards fail often
enough that a bridge is compiled for them, only 1.2\% to 3.6\% of all guards
fail often enough that a bridge is compiled. Also, of all failing guards a few fail extremely often
and most fail rarely. The results emphasize that as most of the guards never
fail it is important to make sure that the successful execution of a guard does
not have unnecessary overhead.

>>>>

It is important to handle guards well, because they are very common operations
in the traces produced by tracing JITs. As we will see later (Figure~\ref{fig:benchmarks})
guards account for about 14\% to 22\% of the
operations before and for about 15\% to 20\% of the operations after optimizing
the traces generated for the different benchmarks used in this paper. An
additional property is that guard failure rates are very uneven. The majority
of guards never fail at all, whereas those that do usually fail extremely
often.

Besides being common, guards have various costs associated with them.
Guards have a runtime cost, they take time to execute. Therefore it is
important to make the on-trace execution of guards as efficient as possible. On
the other hand, guards are possible deoptimization points. The recorded and
compiled path has to be left if a guard fails, returning control to the
interpreter. Therefore guards need enough associated information to enable
rebuilding the interpreter state. The memory overhead of this information
should be kept low. These constraints and trade-offs are what makes the design
and optimization of guards an important and non-trivial aspect of the construction
of a tracing just-in-time compiler.

@@

@@

\begin{figure}
    \include{figures/failing_guards_table}
    \caption{Failing guards, guards with more than 200 failures and guards responsible for 50\% of the failures relative to the total number of guards}
    \label{fig:failing_guards}
\end{figure}

From Figure~\ref{fig:failing_guards} we can see that only a very small amount
of all the guards in the compiled traces ever fail. This amount varies between
2.4\% and 5.7\% of all guards. As can be expected, even fewer guards fail often
enough that a bridge is compiled for them, only 1.2\% to 3.6\% of all guards
fail often enough that a bridge is compiled. Also, of all failing guards a few fail extremely often
and most fail rarely. Reinforcing this notion the figure shows that, depending on the
benchmark, between 0.008\% and 0.225\% of the guards are responsible for 50\%
of the total guards failures.
These results emphasize that as most of the guards never
fail it is important to make sure that the successful execution of a guard does
not have unnecessary overhead.
